Interisland movements of woodland caribou in Newfoundland

### 2020-02-21 paper/manuscript.Rmd
blah blah blah

## Introduction
Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond_2016] 
to avoid predators [@Bergerud_1985] and access islands during calving
(Bergerud et al. 1990). Even for adept swimmers like caribou, the
energetic expenditure of swimming for quadrupedal mammals is
significantly higher than walking or running (Fish 1993), while drowning
is also possible (Miller & Gunn 1985). Despite a large number of coastal and
island caribou herds, only a handful of examples have documented caribou
swimming in the ocean (Table 1).

The Fogo Island archipelago, off the coast of Newfoundland, Canada (49°N, 54°W),
is home to approximately 300 woodland caribou. During routine fieldwork on May
30, 2017, we observed an unmarked adult male caribou swim between Perry and
Eastern Indian Islands (Figure 1), a swim of at least 470 m which took ~9
minutes (~52 m per minute). This observation prompted us to investigate whether swimming behaviour was common among GPS radio-collared caribou (n = 29 adult females) in the population. We identified swimming events as two consecutive GPS locations from an
individual on different islands. This region typically experiences pack ice
during the winter, so caribou may be able to travel between islands by
walking over the ice. We restricted our GPS data to the ice-free period of the
year (April 1 - November 25) to avoid this possibility. Using these criteria, we
identified three adult female caribou that collectively engaged in 86 swimming
events (Figure 2).
<!--Is it better to just state our results here, or give this sort of teaser? Initially I had "86 events by 3 individuals" here, but then we ignore that for a couple paragraphs of context before getting back into it... which seemed weird-->

>>>>

## Introduction
Caribou (*Rangifer tarandus*) are exceptional swimmers. Ample evidence exists
that caribou swim in streams, rivers, and lakes during migration [@Leblond_2016] 
to avoid predators [@Bergerud_1985] and access islands during calving
[@Bergerud_1990]. Even for adept swimmers like caribou, the
energetic expenditure of swimming for quadrupedal mammals is
significantly higher than walking or running (Fish 1993), while drowning
is also possible [@Miller_1986]. Despite a large number of coastal and
island caribou herds, only a handful of examples have documented caribou
swimming in the ocean (Table 1).

The Fogo Island archipelago, off the coast of Newfoundland, Canada (49°N, 54°W),
is home to approximately 300 woodland caribou. During routine fieldwork on May
30, 2017, we observed an unmarked adult male caribou swim between Perry and
Eastern Indian Islands (Figure 1), a swim of at least 470 m which took ~9
minutes (~52 m per minute). This observation prompted us to investigate whether
swimming behaviour was common among GPS radio-collared caribou in the population
(n = 29 adult females: 1240g, GPS 4400M; Lotek Wireless Inc., Newmarket,
Ontario, Canada). We identified swimming events when two consecutive GPS (2-hour
relocation rates) locations from an individual occurred on different islands.
This region typically experiences pack ice during the winter, so caribou may be
able to travel between islands by walking over the ice. We restricted our GPS
data to the ice-free period of the year (April 1 - November 25) to avoid this
possibility. Using these criteria, we identified three adult female caribou that
collectively engaged in 86 swimming events (Figure 2).

<!--Is it better to just state our results here, or give this sort of teaser? Initially I had "86 events by 3 individuals" here, but then we ignore that for a couple paragraphs of context before getting back into it... which seemed weird-->
